Inference of Bifurcations with Differentiable Continuation

### 2020-07-27 docs/article.tex
add draft watermark

\subsection{Similar Works}

\begin{itemize}
    \item some juliacon submissions?
\end{itemize}
\clearpage

@@

\subsection{Optimisation Loop}
We use \texttt{BifurcationKit.jl} \cite{Veltz2019PseudoArcLengthContinuation.jl} to calculate bifurcation diagrams, which requires the rate function $\rates$, its Jacobian $J_\theta$, and an initial guess $u_0$ at an initial parameter $p_0$. Let this algorithm be called by function $\mathrm{getSteadyStates}$ and return the steady states that form the integration region $\Omega$ and a new guess $u_0$ that is the true steady state at $p_0$.

@@

The evaluation of the cost gradient \eqref{eq:gradient}

\\
\begin{algorithm*}[H]
\label{alg:optimisation-loop}
\SetAlgoLined
\textbf{Inputs} Function $F_{\theta}:\mathbb{R}^{N+1}\rightarrow\mathbb{R}^{N}$ for $u\in\mathbb{R}^N$ and $p\in\mathbb{R}$\\ target bifurcations $\mathcal{D}=\{p:p\in[p_{\min},p_{\max}]\}$ and convergence tolerance $\varepsilon$\\
\textbf{Output} Optimised $\theta\in\mathbb{R}^M$ that satisfy targets $\targets$\\
$J_{\theta}\leftarrow\partial F/\partial u$\qquad\,
$\Delta_{\theta}\leftarrow|\mathrm{det}\,J_{\theta}|$\qquad$K_{\theta}\leftarrow\partial^2\Delta_{\theta}/\partial\tangent^2$\\
$\frac{\partial z}{\partial\theta}\leftarrow- \frac{\partial F}{\partial z}^\top
    \left(\,
        \frac{\partial F}{\partial z}\,\frac{\partial F}{\partial z}^\top
    \right)^{-1}
    \frac{\partial F}{\partial\theta}$\\
$u_0\leftarrow\mathrm{rand}\in\mathbb{R}^{N}$\quad\,$\theta\leftarrow\mathrm{rand}\in\mathbb{R}^{M}$\\
$\beta\leftarrow\mathrm{getHyperparameters}(\targets)$\\
\While{$\mathrm{tolerance}(\beta)>\varepsilon$}{
  $\Omega,u_0\leftarrow\mathrm{getSteadyStates}(F_{\theta},J_{\theta},u_0,\beta)$\\
  $\beta\leftarrow\mathrm{updateHyperparameters}(\beta,\Omega,\targets)$\\
  $\partial L\leftarrow\mathrm{costGradient}(\Delta_{\theta},K_{\theta},V_{\theta},\Omega,\targets)$\\
  $\theta\leftarrow\mathrm{updateParameters}(\theta,\partial L,\beta)$
}
\caption{Bifurcation Optimisation Loop}
\end{algorithm*}

>>>>

\subsection{Similar and Related Works}

\begin{itemize}
    \item smooth and match estimator methods \cite{Ranciati2017BayesianParameters}
    \item bifurcation inference using Mixed Integer Nonlinear Programming optimization strategy \cite{Otero-Muras2018Optimization-basedModels}
\end{itemize}
\clearpage

@@

\subsection{Optimisation Loop}
Here provide a high-level view of Algorithm \ref{alg:optimisation-loop}. We use \texttt{BifurcationKit.jl} \cite{Veltz2019PseudoArcLengthContinuation.jl} to calculate bifurcation diagrams, which requires the rate function $\rates$, its Jacobian $J_\theta$, and an initial guess $u_0$ at an initial parameter $p_0$. Let this algorithm be called by function $\mathrm{getSteadyStates}$ and return the steady states that form the integration region $\Omega$ and a new guess $u_0$ that is the true steady state at $p_0$.

@@

The evaluation of the cost gradient \eqref{eq:gradient} is done by $\mathrm{costGradient}$ using \texttt{Zygote.jl} \cite{Innes2018DontPrograms} and requires the following analytical forms: the determinant of the Jacobian $\Det$, its curvature $K_{\theta}$ and the integration region velocity $\frac{\partial z}{\partial\theta}$. It also requires the integration region $\Omega$ and the locations of targets $\targets$. This gradient is used by $\mathrm{updateParameters}$, which can be any gradient-based optimiser such as ADAM or Momentum gradient decent, to update the parameters $\theta$.
\\\\
\begin{algorithm*}[H]
\label{alg:optimisation-loop}
\SetAlgoLined
\textbf{Inputs} Function $F_{\theta}:\mathbb{R}^{N+1}\rightarrow\mathbb{R}^{N}$ for $u\in\mathbb{R}^N$ and $p\in\mathbb{R}$\\ target bifurcations $\mathcal{D}=\{p:p\in[p_{\min},p_{\max}]\}$ and convergence tolerance $\varepsilon$\\
\textbf{Output} Optimised $\theta\in\mathbb{R}^M$ that satisfy targets $\targets$\\
$J_{\theta}\leftarrow\partial F/\partial u$\qquad\,
$\Delta_{\theta}\leftarrow|\mathrm{det}\,J_{\theta}|$\qquad$K_{\theta}\leftarrow\partial^2\Delta_{\theta}/\partial\tangent^2$\\
$\frac{\partial z}{\partial\theta}\leftarrow- \frac{\partial F}{\partial z}^\top
    \left(\,
        \frac{\partial F}{\partial z}\,\frac{\partial F}{\partial z}^\top
    \right)^{-1}
    \frac{\partial F}{\partial\theta}$\\
$u_0\leftarrow\mathrm{rand}\in\mathbb{R}^{N}$\quad\,$\theta\leftarrow\mathrm{rand}\in\mathbb{R}^{M}$\\
$\beta\leftarrow\mathrm{getHyperparameters}(\targets)$\\
\While{$\mathrm{tolerance}(\beta)>\varepsilon$}{
  $\Omega,u_0\leftarrow\mathrm{getSteadyStates}(F_{\theta},J_{\theta},u_0,\beta)$\\
  $\beta\leftarrow\mathrm{updateHyperparameters}(\beta,\Omega,\targets)$\\
  $\partial L\leftarrow\mathrm{costGradient}(\Delta_{\theta},K_{\theta},\frac{\partial z}{\partial\theta},\Omega,\targets)$\\
  $\theta\leftarrow\mathrm{updateParameters}(\theta,\partial L,\beta)$
}
\caption{Bifurcation Optimisation Loop}
\end{algorithm*}
